---
title: "p8105_hw1_irb2118"
output: github_document
date: "2024-09-16"
author: Ravi Brenner
---

# Introduction

This Rmarkdown document is for P8105 - Data Science 1, homework 1.

# Methods

We will be using Rmarkdown (within Rstudio), git, github as well as some packages from the `tidyverse`.

```{r, message=FALSE}
library(tidyverse)
```

The data comes from penguins table in the `palmerpenguins` package:

```{r}
data("penguins", package = "palmerpenguins")
```

# Results

Here is some basic information about the dataset:

The dataset contains information on various species of penguins from different islands, with measurements taken of key variables across several years. There are `r nrow(penguins)` rows and `r ncol(penguins)` columns, with the following variable names:
`r names(penguins)`. The species included are `r unique(penguins$species)` from the islands `r unique(penguins$island)`, and the years range from `r min(penguins$year)` to `r max(penguins$year)`.

The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.

