---
title: "p8105_hw1_irb2118"
output: github_document
date: "Due: 2024-09-21"
---
Ravi Brenner

# Introduction

This Rmarkdown document is for P8105 - Data Science 1, homework 1.

# Methods

We will be using Rmarkdown (within Rstudio), git, github as well as some packages from the `tidyverse`.

```{r, message=FALSE}
library(tidyverse)
```

The data for problem 1 come from penguins table in the `palmerpenguins` package:

```{r}
data("penguins", package = "palmerpenguins")
```

# Problem 1

## Data description
Here is some basic information about the dataset:

The dataset contains information on various species of penguins from different islands, with measurements taken of key variables across several years. There are `r nrow(penguins)` rows and `r ncol(penguins)` columns, with the following variable names:
`r names(penguins)`. The species included are `r unique(penguins$species)` from the islands `r unique(penguins$island)`, and the years range from `r min(penguins$year)` to `r max(penguins$year)`.

The mean flipper length in mm is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.

## Plot
Here is a scatter plot of flipper_length_mm vs. bill_length_mm:
```{r, message = FALSE, warning = FALSE}
ggplot(penguins,
       aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point()

ggsave("species_scatter.png")
```

# Problem 2
Here is a data frame with some special variables, meant to demonstrate how R treats different variable types in practice
```{r}
df_demo <- tibble(
  norm_samp = rnorm(10),
  g0 = if_else(norm_samp > 0, TRUE, FALSE),
  char_vec = c("A","B","C","D","E","F","G","H","I","J is the tenth letter"),
  factor_vec = factor(c("red","green","blue","red","red","blue","red","green","blue","blue"))
)
```

Attempting to take the mean of each variable, we get missing (NA) values for the character vector and factor vector, since  `mean()` only works for numeric, integer, and logical types.
```{r}
mean(pull(df_demo, norm_samp))
mean(pull(df_demo, g0))
mean(pull(df_demo, char_vec))
mean(pull(df_demo, factor_vec))
```

Attempting to convert the logical, character, and factor variables to numeric. This  works for the logical value (1 for `TRUE` and 0 for `FALSE`) and for the factor variable (R converted the 3 factors to numeric values 1, 2, and 3). There isn't anything clear to do with the character vector, so R returns a vector of NAs. When we tried to take the mean, R implicitly converted the logical vector to numeric and returned that mean. For the character vector, it returned NA (since each value was NA when converted to numeric). For the factor vector, the `mean()` function also returned NA, since the numeric values returned by `as.numeric()` are really just ordinal labels for the 3 factors and they don't have any intrinsic value, so it doesn't make sense to return a mean for those values.
```{r, eval=FALSE}
as.numeric(pull(df_demo, g0))
as.numeric(pull(df_demo, char_vec))
as.numeric(pull(df_demo, factor_vec))
```

# Conclusion
This homework demonstrated some basic ways to use R to manipulate dataframes and variables, using Rmarkdown inline text and code chunks.





